---
title: "findMatch"
author: 
- name: "Sigbert Klinke" 
  email: sigbert@hu-berlin.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
vignette: > 
  %\VignetteIndexEntry{plot.matrix} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,  include=FALSE}
library(findMatch)
```

# Creating some random data sets

## For one point in time 
A small dataset

```{r}
d <- generateTestData(6)
head(d)
```

A small dataset with an additional variable `points`

```{r}
d <- generateTestData(6, points=function(n) { sample(0:20, size=n, replace=TRUE)} )
head(d)
```

A small dataset with overwriting the `birthplace` with a vector

```{r}
d <- generateTestData(6, birthplace=c("Berlin", "Hamburg", "Köln", "München"))
head(d)
```

A small dataset with overwriting the `birthplace` with a function

```{r}
d <- generateTestData(6, birthplace=function(n) {
                            sample(c("Berlin", "Hamburg", "Köln", "München"), 
                                   size=n, replace=TRUE, 
                                   prob=c(3520031, 1787408, 1060582, 1450381))
                      }) 
head(d)
```

## For more two points in time 

Two small data sets with 6 observations at t1 and 4 observations at t2 without overlapping observations

```{r}
d <- generateTestData(c(6, 4))
str(d)
```

For two data sets with 6 observations only in t1, 4 observations only in t2 and 5 observations only in t1 and t2 
you have to construct a list of vectors. Each vector has as first entry the number of observations and as further entries the number of the timepoints these observations should be. For example

* `c(6, 1)` means 6 observations only at t1 or
* `c(4, 2)` means 4 observations only at t2 or
* `c(5, 1, 2)` means 5 observations at t1 and t2.

This creates two data frames with an appropriate observation structure

```{r}
# t1: 6+5=11 observations
# t2: 4+5=9 observations
n <- list(c(6, 1), c(4, 2), c(5, 1, 2))
str(n)
d <- generateTestData(n)
str(d)
```

## For more than two points in time 

Three data frames with 

* 6 observations only in t1, 
* 4 observations only in t2,
* 2 observations only in t3,
* 5 observations in t1 and t2,
* 8 observations in t1 and t3,
* 3 observations in t2 and t3 and 
* 7 observations in t1, t2 and t3.

```{r}
# t1: 6+5+8+7=26 observations
# t2: 4+5+3+7=19 observations
# t3: 2+8+3+7=20 observations
n <- list(c(6, 1), c(4, 2), c(2, 3), c(5, 1, 2), c(8, 1, 3), c(3, 2, 3), c(7, 1, 2, 3))
str(n)
d <- generateTestData(n)
str(d)
```

# Find matches between two points in time

At first we generate two test data sets and then match on the `code` variables using the Levenshtein distance. We most likely found 5 matches.

```{r}
n <- list(c(6, 1), c(4, 2), c(5, 1, 2))
data  <- generateTestData(n)
vars  <-  c("code", "code")
match <- findMatch(data, vars)
# 
summary(match)
# 
head(match)
```

The `summary` tells us that we found 5 perfect matches with Levensthein distances of zero. 

Each line in `head` should read as follows:

* `line.1`: observation number in data frame 1
* `line.2`: observation number in data frame 2
* `uid.1`, `uid.2`: unique ids over all data sets for observation, if not given then `dataset:lineno` is used
* `idn.0.ZDV`: common code created from `vars` 
* `idn.1.code`, `idn.2.code`: codes compared with
* `leven.1`, `leven.2`: Levensthein distances between common code and codes in the two data frames
* `leven.V3`: sum of `leven.1` and `leven.2`



We may allow for a larger Levenshtein distance and find two more possible matches, but they are not exact

```{r}
set.seed(0)
n <- list(c(6, 1), c(4, 2), c(5, 1, 2))
data  <- generateTestData(n)
vars  <-  c("code", "code")
match <- findMatch(data, vars, dmax=5)
# 
summary(match)
# 
head(match)
```

# Find matches between more than two points in time


